\textit{Supporting Information for}: Geometric models reveal behavioral and neural signatures of transforming naturalistic experiences into episodic memories

### 2020-09-04 paper/supporting_information.tex
minor changes to supp

\section*{Additional details about topic modeling methods and results}
\subsection*{Optimizing topic model parameters}
In order to create accurate episode and recall models, we used an optimization method that was driven by our ability to explain hand-annotated memory performance metrics collected by \cite{ChenEtal17}.  In an earlier variant of our study~\citep{HeusMann18}, we used a grid search to compute the $\omega$ (episode sliding window duration, in scenes), $\rho$ (recall sliding window duration, in sentences), and $K$ (number of topics) that satisfied
\[
\argmax_{\omega, \rho, K} \left[\mathrm{corr}\left(\mathrm{corr}\left(\mu\left(\omega, \rho, K\right), \nu\left(\omega, \rho, K\right)\right), \theta\right)\right],
\]
where $\mathrm{corr}(\mu, \nu)$ is the per-participant correlation between the temporal correlation matrices of the episode ($\mu$) and recall ($\nu$) topic proportions matrices, and $\theta$ is the per-participant hand-counted number of recalled scenes.  We searched over a grid of pre-specified values for each of these parameters; the resulting correlations are displayed in Figure~\ref{fig:paramsearch}.  The optimal parameters were $\omega = 50$, $\rho = 10$, and $K = 100$.  In our current paper we made a number of improvements to how we preprocessed text and fit topic models (see \textit{Methods}), but we carried the same optimal parameters forward from \cite{HeusMann18} without performing any additional optimization.

\begin{figure}[b]
\centering
\includegraphics[width=1\textwidth]{figs/parameter_search}
\caption{\small \textbf{Optimizing topic model parameters.}  We performed a grid search over episode sliding window length ($\omega \in \{5, 10, 25, 50, 100 \}$), recall sliding window length ($\rho \in \{5, 10, 25, 50, 100 \}$), and number of topics ($K \in \{5, 10, 25, 50, 100 \}$).  The reported correlations are between per-subject episode-recall correlations and per-participant hand-counted numbers of recalled scenes.}
\label{fig:paramsearch}
\end{figure}

@@

\begin{figure}[p]
\centering
\includegraphics[width=0.65\textwidth]{figs/topic_words}
\caption{\small \textbf{Topics discovered in \textit{Sherlock}.} We applied a topic model to hand-annotated information about 1000 scenes spanning the 48 minute episode.  We identified 32 unique topics with non-zero weights (we used $K=100$ topics to fit the model).  Each topic comprises a distribution of weights over all words in the vocabulary.  For each topic, we show the words with the 10 largest weights, along with a suggested description of the topic.}
\label{fig:topics}
\end{figure}
%\FloatBarrier

\subsection*{Feature importance analyses}
To determine the contribution of each feature to the temporal structure of the episode topic proportions matrix, we conducted a ``leave one out'' analysis.  Specifically, we compared the original episode's topic proportions matrix (created using all hand-annotated features from the 998 manually identified scenes spanning the \textit{Sherlock} episode; see \textit{Methods} for a full list of features) with topic proportions matrices created using all but one type of feature.  For each impoverished topic proportions matrix, we computed the timepoint-by-timepoint correlation matrix, and correlated the proximal diagonals from the upper right triangle with those from the temporal correlation matrix of the feature-complete matrix (for details on how we isolated proximal temporal correlations, see \textit{Methods}).  Observing a lower correlation between an impoverished matrix (with a particular feature removed) and the feature-complete matrix would suggest that the held-out feature contributed more prominently to the full episode topic proportion matrix's temporal structure.  We found that hand-annotated narrative details played the greatest roll in determining the temporal structure of the episode, whereas the name of the character(s) in focus for each shot contributed least (Fig.~\ref{fig:feature-importance}A).

@@

Next, we sought to determine which annotated features contributed aspects of the episode's temporal structure that were preserved in participants' later recalls.  Specifically, we computed the timepoint-by-timepoint correlation matrix of the episode's topic proportions matrix, and correlated the proximal diagonals from its upper triangle with those from the timepoint-by-timepoint correlation matrices for each participant's recall topic proportions matrices (stretched via linear interpolation to have the same number of timepoints as the episode topic proportions matrix).  This yielded a single correlation coefficient for each participant.  We then repeated this analysis with each annotated feature held out in turn.  Observing a lower correlation between the episode and recall topic proportions matrices (constructed in the absence of a given feature) would indicate that participants utilize changes in that feature's content to discriminate between sections of the episode when organizing their recalls.  We found that hand-annotated narrative details were the most heavily utilized feature, whereas changes in the text present on-screen, the indoor/outdoor distinction between scenes, the camera angle, the names of the various characters on screen, and the location in which the scene took place tended not to impact participants' recall structures (i.e., removing those features resulted in a \textit{greater} episode-recall correlation than including them; Fig.~\ref{fig:feature-importance}B).

@@

\newpage
\renewcommand{\refname}{Supplemental references}
% \bibliography{../../CDL-bibliography/memlab}
\bibliography{CDL-bibliography/memlab}

>>>>

\section*{Additional details about topic modeling methods and results}
\subsection*{Optimizing topic model parameters}
In order to create accurate episode and recall models, we used an optimization method that was driven by our ability to explain hand-annotated memory performance metrics collected by \cite{ChenEtal17}.  In an earlier variant of our study~\citep{HeusMann18}, we used a grid search to compute the $\omega$ (episode sliding window duration, in scenes), $\rho$ (recall sliding window duration, in sentences), and $K$ (number of topics) that satisfied
\[
\argmax_{\omega, \rho, K} \left[\mathrm{corr}\left(\mathrm{corr}\left(\mu\left(\omega, \rho, K\right), \nu\left(\omega, \rho, K\right)\right), \theta\right)\right],
\]
where $\mathrm{corr}(\mu, \nu)$ is the per-participant correlation between the episode ($\mu$) and recall ($\nu$) topic proportions matrices, and $\theta$ is the per-participant hand-counted number of recalled scenes.  We searched over a grid of pre-specified values for each of these parameters; the resulting correlations are displayed in Figure~\ref{fig:paramsearch}.  The optimal parameters were $\omega = 50$, $\rho = 10$, and $K = 100$.  In our current paper we made a number of improvements to how we preprocessed text and fit topic models (see \textit{Methods}), but we carried the same optimal parameters forward from \cite{HeusMann18} without performing any additional optimization.

\begin{figure}[b]
\centering
\includegraphics[width=1\textwidth]{figs/parameter_search}
\caption{\small \textbf{Optimizing topic model parameters.}  We performed a grid search over episode sliding window length ($\omega \in \{5, 10, 25, 50, 100 \}$), recall sliding window length ($\rho \in \{5, 10, 25, 50, 100 \}$), and number of topics ($K \in \{5, 10, 25, 50, 100 \}$).  The reported correlations are between per-participant episode-recall correlations and per-participant hand-counted numbers of recalled scenes.}
\label{fig:paramsearch}
\end{figure}

@@

\begin{figure}[p]
\centering
\includegraphics[width=0.65\textwidth]{figs/topic_words}
\caption{\small \textbf{Topics discovered in \textit{Sherlock}.} We applied a topic model to hand-annotated information about 1000 scenes spanning the 48-minute episode.  We identified 32 unique topics with non-zero weights (we used $K=100$ topics to fit the model).  Each topic comprises a distribution of weights over all words in the vocabulary.  For each topic, we show the words with the 10 largest weights, along with a suggested description of the topic.}
\label{fig:topics}
\end{figure}
%\FloatBarrier

\subsection*{Feature importance analyses}
To determine the contribution of each feature to the temporal structure of the episode topic proportions matrix, we conducted a ``leave one out'' analysis.  Specifically, we compared the original episode's topic proportions matrix (created using all hand-annotated features from the 998 manually identified scenes spanning the \textit{Sherlock} episode; see \textit{Methods} for a full list of features) with topic proportions matrices created using all but one type of feature.  For each impoverished topic proportions matrix, we computed the timepoint-by-timepoint correlation matrix, and correlated the proximal diagonals from the upper triangle with those from the temporal correlation matrix of the feature-complete matrix (for details on how we isolated proximal temporal correlations, see \textit{Methods}).  Observing a lower correlation between an impoverished matrix (with a particular feature removed) and the feature-complete matrix would suggest that the held-out feature contributed more prominently to the full episode topic proportion matrix's temporal structure.  We found that hand-annotated narrative details played the greatest roll in determining the temporal structure of the episode, whereas the name of the character(s) in focus for each shot contributed least (Fig.~\ref{fig:feature-importance}A).

@@

Next, we sought to determine which annotated features contributed aspects of the episode's temporal structure that were preserved in participants' later recalls.  Specifically, we computed the timepoint-by-timepoint correlation matrix of the episode's topic proportions matrix, and correlated the proximal diagonals from its upper triangle with those from the timepoint-by-timepoint correlation matrices for each participant's recall topic proportions matrices (stretched via linear interpolation to have the same number of timepoints as the episode topic proportions matrix).  This yielded a single correlation coefficient for each participant.  We then repeated this analysis with each annotated feature held out in turn.  Observing a lower correlation between the episode and recall topic proportions matrices (constructed in the absence of a given feature) would indicate that participants utilize changes in that feature's content to discriminate between sections of the episode when organizing their recalls.  We found that hand-annotated narrative details were the most heavily utilized feature, whereas changes in the text present on-screen, the indoor/outdoor distinction between shots, the camera angle, the names of the various characters on screen, and the shot's location tended not to impact participants' recall structures (i.e., removing those features resulted in a \textit{greater} episode-recall correlation than including them; Fig.~\ref{fig:feature-importance}B).

@@

\newpage
\renewcommand{\refname}{Supplemental references}
\bibliography{CDL-bibliography/memlab}
